---
title: "Lab 6"
font-family: 'Corbel'
author: Patrick Sinclair, Kieran Yuen
output: github_document
---
```{r}
setwd("/Users/kieranyuen/Documents/Econometrics/Econometrics")
load("acs2017_ny_data copy.RData")
attach(acs2017_ny)
detach()
```

## Lab 6

Anyway, on to the lab. This week we move to logit and probit models. These are suited for when the dependent y variable takes values of just 0 or 1.

We will look at labor force participation, which is an important aspect of the economy -- who chooses to work (where those who are unemployed but looking for work are considered part of the labor force).


The main differences from lm() are that the call now has a g in it, glm(), for Generalized Linear Model; and that it includes the bit about family = binomial.

*Before you rush off to estimate that* just note that there is some preliminary work to be done (such as creating the data to use).

```{r eval=FALSE}
acs2017_ny$LABFORCE <- as.factor(acs2017_ny$LABFORCE)
levels(acs2017_ny$LABFORCE) <- c("NA","Not in LF","in LF")
acs2017_ny$MARST <- as.factor(acs2017_ny$MARST)
levels(acs2017_ny$MARST) <- c("married spouse present","married spouse absent","separated","divorced","widowed","never married")
```
What is the difference between "NA" as label and Not in the Labor Force? Make sure you understand. (Hint, look at ages in each group).
NA - not able to participate in the labor force
Not in LF - means they can participate in the labor force but they choose not to



In general it is a good idea to check summary stats before doing fancier models. What fraction of people, say, 55-65, are in the labor force? What about other age ranges? What would you guess are other important predictors? For example,

```{r eval=FALSE}
workers <- (acs2017_ny$LABFORCE =="in LF")
dat_workers <- subset(acs2017_ny,workers)
```



```{r eval=FALSE}
acs2017_ny$age_bands <- cut(acs2017_ny$AGE,breaks=c(0,15, 25,35,45,55,65,100))
table(acs2017_ny$age_bands,acs2017_ny$LABFORCE)
```
(although kable package could do nicer, if you want to play)

# Model: 25-55
```{r eval=FALSE}
pick_use1 <- (acs2017_ny$AGE >25) & (acs2017_ny$AGE <= 55)
dat_use1 <- subset(acs2017_ny, pick_use1)
dat_use1$LABFORCE <- droplevels(dat_use1$LABFORCE) # actually not necessary since logit is smart enough to drop unused levels, but helps my personal sense of order
```


```{r eval=FALSE}
model_logit1 <- glm(LABFORCE ~ AGE + female + AfAm + Asian + race_oth + Hispanic 
            + educ_hs + educ_somecoll + educ_college + educ_advdeg 
            + MARST,
            family = binomial, data = dat_use1)
summary(model_logit1)

plot(LABFORCE ~ jitter(AGE, factor = 1), pch = 16, col = rgb(0, 0.5, 0, alpha = 0.2), ylim = c(0,1), data = dat_use1)

#logitpredict1 <- data.frame(AGE = 25:35)
#dat_use1$yhat <- predict(model_logit1, newdata = dat_use1)
#lines(yhat ~ AGE, data = dat_use1, col = "Black")
#legend("topleft", c("Mid"), lty = 1, col=c("Black"), bty = "n")
#max(dat_use1$yhat) 
#median(dat_use1$yhat)
#print(dat_use1$yhat)

```
























# Start of Kieran's work
What other X variables might you add? Maybe some interactions? LOTS OF INTERACTIONS? What other subsets? What changes about results with different subsets?

We would like to use FAMSIZE to see how family size affects a person's probabilities of being in the labor force. We decided to break up the family size column into three groups (# of family members): small (0-3), medium (4-6), large (7-29). 

Let's first take a bird's eye view of the acs2017 data to see how many of our three Family size groups are in the Labor Force.

Results: Looking at the overview table of our data we can see that a majority of people of all ages and family sizes participate in the labor force. This is to be expected as most people do need to make a living and bring in income to survive. But we see there is also a good proportion of those who are not in the labor force, which allows us to be able to do a regression analysis on the difference.
```{r eval=FALSE}
acs2017_ny$SizeofFam <- cut(acs2017_ny$FAMSIZE,breaks=c(0,3,6,29))
table(acs2017_ny$LABFORCE,acs2017_ny$age_bands,acs2017_ny$SizeofFam)
```


Next, let's make separate columns in the acs2017 data for each of the three (3) family sizes similar to how there are five (5) columns for education (educ_nohs, educ_hs, etc....).
```{r eval=FALSE}
acs2017_ny$SmallFamily <- ((acs2017_ny$FAMSIZE >= 0) & (acs2017_ny$FAMSIZE <= 3))
acs2017_ny$MediumFamily <- ((acs2017_ny$FAMSIZE >= 4) & (acs2017_ny$FAMSIZE <= 6))
acs2017_ny$LargeFamily <- ((acs2017_ny$FAMSIZE >= 7) & (acs2017_ny$FAMSIZE <= 29))
```

# Logit: Fam Size ONLY
Next, let's run the logit regression to see how the family sizes affect whether someone is in the labor force or not. 

Expectation: We are expecting that one of the three (3) variables will drop off as that is necessary so that the variables can be compared to something. also, we were thinking that the larger the size of the family the less likely the labor participation rate would be, but the counter to that is that those larger families have more mouths to feed and thus would increase their likelihood to be in the labor force. So it is hard to estimate which of the three (3) family sizes will show to have the largest effect on labor participation rates.

Result: So it appears that *medium* families have the highest probability of being in the labor force. Next would be *small* families and last is *large* families which have the lowest probabilities of being in the labor force.
```{r eval=FALSE}
model_logit2 <- glm(LABFORCE ~ SmallFamily + MediumFamily + LargeFamily,
            family = binomial, data = dat_use1)
summary(model_logit2)

model_logit3 <- glm(LABFORCE ~  MediumFamily + LargeFamily + SmallFamily,
            family = binomial, data = dat_use1)
summary(model_logit3)

model_logit4 <- glm(LABFORCE ~  LargeFamily + SmallFamily + MediumFamily,
            family = binomial, data = dat_use1)
summary(model_logit4)
```

# Logit: All Variables
Next, let's add in all the other variables to see how that affects our probabilities.

Result: Medium families show the highest probability. Followed by small families. Last is large families. All three family sizes are statistically significant at the 0.05 level.
```{r eval=FALSE}
model_logit5 <- glm(LABFORCE ~ AGE + female + SmallFamily + MediumFamily + LargeFamily,
            family = binomial, data = dat_use1)
summary(model_logit5)

model_logit6 <- glm(LABFORCE ~ AGE + female + MediumFamily + LargeFamily + SmallFamily,
            family = binomial, data = dat_use1)
summary(model_logit6)

model_logit7 <- glm(LABFORCE ~ AGE + female + LargeFamily + SmallFamily + MediumFamily,
            family = binomial, data = dat_use1)
summary(model_logit7)
```

# Logit: All Variables (Interactions)
```{r eval=FALSE}
model_logit8 <- glm(LABFORCE ~ AGE + female + MediumFamily + LargeFamily + SmallFamily + I(MediumFamily*female) + I(LargeFamily*female) + I(SmallFamily*female) + I(female*AGE),
            family = binomial, data = dat_use1)
summary(model_logit8)
tab_model(model_logit8, p.style = "numeric_stars", transform = NULL)
plot_model(model_logit8)+theme_blank()

model_logit9 <- glm(LABFORCE ~ AGE + female + MediumFamily + LargeFamily + SmallFamily + I(MediumFamily*female) + I(LargeFamily*female) + I(SmallFamily*female) + I(female*AGE),
            family = binomial, data = dat_use1)
summary(model_logit9)

model_logit10 <- glm(LABFORCE ~ AGE + female + LargeFamily + SmallFamily + MediumFamily + I(LargeFamily*AGE) + I(SmallFamily*AGE) + I(MediumFamily*AGE),
            family = binomial, data = dat_use1)
summary(model_logit10)
```

How do we do the zero-to-one plots?

```{r eval=FALSE}
#graveyard <- (( >= 0) & (DEPARTS <= 459))
#morning <- ((DEPARTS >= 500) & (DEPARTS <=930))
#daytime <- ((DEPARTS >= 931) & (DEPARTS <=1700))
#evening <- ((DEPARTS >= 1701) & (DEPARTS <=2359))
#summary(graveyard)
#summary(morning)
#summary(daytime)
#summary(evening)
#table(morning)
```




# For homework, I will ask for predicted values so you can start to figure out how to get those.

# Do the X variables have the expected signs and patterns of significance? Explain if there is a plausible causal link from X variables to Y and not the reverse. Explain your results, giving details about the estimation, some predicted values, and providing any relevant graphics. Impress.


# Start of Probit Models

Also estimate a probit model (details in Lecture Notes) and OLS, with the same X and Y variables. Compare the results, such as coefficients and predicted values. If you're eager, try to split the sample into training and test data, then compare which model predicts better in the sample that it hasn't seen yet.

Logit models are very very commonly used in many models. But on the other hand there are people that say it says too much. Sometimes bland "OLS" might be better

# Notes
```{r eval=FALSE}
#How do OLS and logit compare?

#OLS is just creating a line. 

#Our y variable can only be equal to 0 or 1, this can possibly be a problem for some situations.

#With a logit, it is strained to produce a value between 0 and 1.

#difference between logit and probit? Usually very minor.....

#think of it that "not one model is right," think of it that we need to use all these models. To what extent am i confident on using these as a basis for a prediction?

#I can use a logit, probit and OLS model. And then amalgamate all three together

#it depends on the range of your x variables

#greater than 1 means there are probably other factors

#generally, we can easily interpret the sign of a coefficient, but the actual number of the coefficient is not easily interpreted as many 

#what should we do with dropped variables? It has to drop to one variable, so that people who are 

#the amount people spend on their utility is probably gonna be related to their income
#if i make a lot of money, i might move somewhere that is nicer that probably will have higher utility costs

#logit case - the Y can only take two possible answers: is the person in the labor force or not? Does someone have health insurance? We cannot do this with marital status because someone would be unmarried now but they were married once before.

#How do people choose, what type of commute they use to work?
#Logit: did people take public transportation or not?
#subdivide: if it is public transit, what form of public transit?
#COVID-19: Did the person work from home or not?
```


## Appendix

To clear up some of the definitions,
```{r eval=FALSE}
acs2017_ny$EMPSTAT <- as.factor(acs2017_ny$EMPSTAT)
levels(acs2017_ny$EMPSTAT) <- c("NA","Employed","Unemployed","Not in LF")
acs2017_ny$LABFORCE <- as.factor(acs2017_ny$LABFORCE)
levels(acs2017_ny$LABFORCE) <- c("NA","Not in LF","in LF")
acs2017_ny$CLASSWKR <- as.factor(acs2017_ny$CLASSWKR)
levels(acs2017_ny$CLASSWKR) <- c("NA","self employed","work for wages")
acs2017_ny$WKSWORK2 <- as.factor(acs2017_ny$WKSWORK2)
levels(acs2017_ny$WKSWORK2) <- c("NA","1-13 wks","14-26 wks","27-39 wks","40-47 wks","48-49 wks","50-52 wks")
# although note that making this a factor breaks some earlier code where we used (WKSWORK2 > 4) so you might not want to run that code or else change to WKSWORK2_factor <- as.factor(WKSWORK2). Which is arguably better for various other reasons.
# these help clarify how these definitions work together
table(acs2017_ny$EMPSTAT,acs2017_ny$LABFORCE)
table(acs2017_ny$EMPSTAT,acs2017_ny$CLASSWKR)
```


